---
title: Code Review by Santiago Rodriquez
subtitle: converted to quarto doc by Libby Heeren
format: html
---

```{r}
# Set up box
box::use(
  here[here],
  bench[mark, press],
  glue[glue]
)

# Source the functions file to load all functions
source(here("R/01-functions.R"))

# Run time checks - only once I've created a revised file
# Need to update file locations in here()

# pdf(NULL)
# # new_runtime = system.time(source(here("original", "code_review_revised.R")))
# invisible(dev.off())
# 
# pdf(NULL)
# new_size = object.size(source(here("original", "code_review.R")))
# invisible(dev.off())
```

# Functions
## get_prob_matrix
#### Initiating empty matrices  

One of the steps in the get_prob_matrix function creates an empty matrix. Depending on how big the matrix is there are more efficient ways to do this. For instance, the `allocMatrix` function in the `matrixStats` library is very fast, but it would introduce a dependency.

#### Error handling

Testing different values of the argument size: 40, 1, 2, 0. Essentially testing how the function handles weird values of size.

::: {style="color:purple;"}
I was thinking, since it will be a Shiny app, I could just set limits on the inputs so that people could only input size as an integer between, say, 13 and 60.  
:::

```{r}
# Explanation from Lib for blog: This code is saying, "try to make a prob matrix of a certain size and then see what the class of the resulting object is - is "matrix" in that list of classes? TRUE or FALSE?" You can run just the try call and see that it returns either a list of classes or an error

"matrix" %in% try(class(get_prob_matrix(size = 40, prob_vector = get_prob_vector(20))))
# TRUE

"matrix" %in% try(class(get_prob_matrix(size = 1, prob_vector = get_prob_vector(20))))
# FALSE

"matrix" %in% try(class(get_prob_matrix(size = 2, prob_vector = get_prob_vector(20))))
# TRUE

"matrix" %in% try(class(get_prob_matrix(size = 0, prob_vector = get_prob_vector(20))))
# FALSE
```
I don’t know how likely it is that the size value is 0 or 1 but there are instances where the function fails. 

::: {style="color:purple;"}
So, size 0, 1, and 2 won't be in the bounds of the Shiny input, but this is a super great way to test that I've NEVER SEEN BEFORE! If I was creating a package from this, I'd 100% want to make sure I had errors and messages handled correctly! (Could I make this a package someday???) This makes me incredibly happy. Learn-o-meter is at 2. (The first count on the Learn-o-meter was from Santiago being the first person to EVER explain why `{box}` is useful in a meaningful (to me) way by comparing it to `from <library> import <function set>` in python.) Also, note for future, Jadey says to look into the `{cli}` package for error handling!
:::

#### if vs ifelse

From the documentation of the `ifelse` function:  

> Further note that if(test) yes else no is much more efficient and often much preferable to ifelse(test, yes, no) whenever test is a simple true/false result, i.e., when length(test) == 1.  

If size is always a scalar (length 1) not an array (length n) then if(test) yes else no will be faster, as can be seen below.  

```{r}
tmp = 40

# mark() is from bench package
mark(
  ifelse(tmp %% 2 == 0, tmp/2, (tmp+1)/2),
  if(tmp %% 2 == 0) tmp/2 else (tmp+1)/2,
  iterations = 5000,
  time_unit = "ns"
)

rm(tmp)
```

::: {style="color:purple;"}
Mind blown 🤯 I seriously hadn't ever considered this and **definitely** hadn't read that in the documentation. That part in the docs about constructing things like this also was new to me: `tmp <- yes; tmp[!test] <- no[!test]; tmp` where yes and no are return values and test is a test expression. Learn-o-meter is at 3.
:::
```{r}
# From Libby: walking through the other example of structure from the ifelse doc

val <- 4

# val[!val < 3] will evaluate to val if val is NOT less than three
# val[!val < 3] will evaluate to numeric(0) if val IS less than three
# 4 is not less than 3, so the code below evaluates to val <- 2
# if val was 1 it would be numeric(0) <- numeric(0) and val would not be changed

val[!val < 3] <- 2[!val < 3] # code runs despite an RStudio error marker

val # val is now 2

rm(val)
```


#### Simplyfing logic

It appears the goal of `quad_size <- if(size %% 2 == 0) size/2 else (size+1)/2` is to ensure that quad_size is always even. If so, then perhaps there’s a simpler way to achieve the same thing. Let’s test some logic below.

The below code tests different values of size: 40, 39, 100, an 99. Two approaches are tested. The first is the updated if(test) logic from above and the other is ceiling(size/2). This is simpler because it doesn’t reference size multiple times as the if(test) size/2 else (size+1)/2 code does.

```{r}
# press() is from bench package
press(
  tmp = c(40, 39, 100, 99),
  {
    mark(
      if(tmp %% 2 == 0) tmp/2 else (tmp+1)/2,
      ceiling(tmp/2),
      iterations = 1000,
      time_unit = "ns"
    )
  }
)
```
As suspected, the ceiling() function is faster than if(test). Now, let’s confirm the two functions return the same values.

```{r}
for(i in c(40, 39, 100, 99, 99.5)){
  print(glue(
    "Size = {i}", " ",
    "Original = {if(i %% 2 == 0) i/2 else (i+1)/2}", " ",
    "Test = {ceiling(i/2)}"
  ))
  
  # house cleaning
  rm(i)
}
```
In conclusion, the ceiling logic is faster and can handle non-integer size values!

::: {style="color:purple;"}
I will need to make sure my Shiny input is integer only for sure. Using `ceiling` makes perfect sense and this is a great example of why an extra set of eyes is so important when you're too deep in a project to see simple fixes easily!
:::

#### DRY - don’t repeat yourself

I noticed that the code_review file had a `circuits` variable that used the same logic as the `quad_size` variable in this function, and that more than one function used this circuits/quad_size logic. So why not make `quad_size` a function argument vs re-calculating `circuits`.

::: {style="color:purple;"}
Definitely me being lazy :D I even have a note in my process documentation that says I should do this but I'm too lazy at the moment hahaha! I will definitely fix this in the final version.
:::

#### Final speed test
In this test, we’ll compare the original function vs the updated function  

```{r}
# From Libby: The two different functions get defined, then tested to see if they
# output the same results
func1 <- function(size, prob_vector){

  # Calculate quad size same way as circuits
  quad_size <- ifelse(size %% 2 == 0, size/2, (size+1)/2)

  # Create empty matrix for the quad
  M <- matrix(0, nrow = quad_size, ncol = quad_size)

  # For loop to assign prob_vector to correct cells in quadrant
  for (i in 1:quad_size){

    M[i, i:quad_size] <- prob_vector[i]
    M[i:quad_size, i] <- prob_vector[i]
  }

  # if size is even, we need to mirror the entire quad
  if(size %% 2 == 0){
    # mirror horizontally and column bind
    M_right <- pracma::fliplr(M) # fliplr means flip left-right
    M <- cbind(M, M_right)

    # then mirror vertically and row bind
    M_down <- pracma::flipud(M) # flipud means flip up-down
    M <- rbind(M, M_down)

  }else{ # if size is odd, then we leave the last col and row out of mirroring
    # mirror all but last col horizontally and col bind
    M_right <- pracma::fliplr(M[ , 1:(quad_size-1)]) # fliplr means flip left-right
    M <- cbind(M, M_right)

    # then mirror all but last row vertically and row bind
    M_down <- pracma::flipud(M[1:(quad_size-1), ]) # flipud means flip up-down
    M <- rbind(M, M_down)

  }

  return(M)
}

func2 <- function(size, quad_size, prob_vector){
  # Create empty matrix for the quad
  M <- matrix(0, nrow = quad_size, ncol = quad_size)

  # For loop to assign prob_vector to correct cells in quadrant
  for (i in seq_len(quad_size)){

    M[i, i:quad_size] <- prob_vector[i]
    M[i:quad_size, i] <- prob_vector[i]
  }

  # if size is even, we need to mirror the entire quad
  if(size %% 2 == 0){
    # mirror horizontally and column bind
    M_right <- pracma::fliplr(M) # fliplr means flip left-right
    M <- cbind(M, M_right)

    # then mirror vertically and row bind
    M_down <- pracma::flipud(M) # flipud means flip up-down
    M <- rbind(M, M_down)

  }else{ # if size is odd, then we leave the last col and row out of mirroring
    # mirror all but last col horizontally and col bind
    M_right <- pracma::fliplr(M[ , seq_len(quad_size-1)]) # fliplr means flip left-right
    M <- cbind(M, M_right)

    # then mirror all but last row vertically and row bind
    M_down <- pracma::flipud(M[seq_len(quad_size-1), ]) # flipud means flip up-down
    M <- rbind(M, M_down)

  }

  return(M)
}

# size
tmp = 100
# circuits
tmp2 = ceiling(tmp/2)
# prob_vector
tmp3 = get_prob_vector(tmp2)


# test match rate
tmpMatchRate = sum(func1(tmp, tmp3) == func2(tmp, tmp2, tmp3)) / length(func1(tmp, tmp3))
glue("func1 output matches func2 output {round(tmpMatchRate,2)*100}%")
# func1 output matches func2 output 100%
```

```{r}
# speed test
mark(
  func1(tmp, tmp3),
  func2(tmp, tmp2, tmp3),
  iterations = 1000,
  time_unit = "ns"
)

# house cleaning
rm(func1, func2, tmp, tmp2, tmp3, tmpMatchRate)
```

Final modifications slightly reduced how long the function takes to run, simplified things via adding a functional argument, and added seq_len where appropriate.  

::: {style="color:purple;"}
I'm going to bump the Learn-o-meter to 4 because, even if my silly little app will never really need to be blazing fast, these are invaluable skills to carry forward with me into all other work I do. I'm definitely making all of these changes!
:::

## get_color_vector
#### Seed

Since the functions make use of sample it might be a good idea to set a seed so that results can be replicated. I added an argument to the function called seed. The default value is 123. However, if that’s not desired, then you could do seed = round(runif(1, 100, 10000)) to randomly assign the seed. This allows users to replicate results (assuming the seed is saved somewhere) but generates different pictures every time.

::: {style="color:purple;"}
This is a feature on my list of possible upgrades! I haven't thought of a way to integrate capturing the seed in the session info yet, but I'd like to be able to do that at some point. I like the ethereal nature of the pieces, but I do think it would be convenient to be able to recreate one I really liked. Maybe I could have a section of the app that was just for recreations. It would just ask for a "plot ID" or something, which would be the seed, and I could make sure that the seed got output somewhere obvious for each plot so the user could capture it. Hmmm. I will need to think on it.
:::

#### Testing changes
Similar to the get_prob_matrix function above what if we replace all of the `ifelse` functions with if(test)? Additionally, the `1:nrow()` and `1:ncol()` functions were replaced with `seq_len(nrow())` and `seq_len(ncol())`. I’ll create another function `func3` to showcase faster alternatives to `unique(matching[which(matching != "")])`.

```{r}
# Do the set-up to be able to run the functions (define size and colors)
tmp1 = 100
tmp2 <- "#EDEFEE"
if (tmp2 == "#EDEFEE"){ # If background is paper white
  tmp3 <- c(#"#EDEFEE", # Paper
    "#1A8BB3", # Teal - no longer teal, just bright blue
    "#0950AE", # Dark blue
    "#4DACE5", # Light blue
    "#126DDB", # Blue
    "#E48DC4", # Pink
    "#ABA9E8", # Light purple
    "#872791", # Purple
    "#6D1617", # Dark red
    "#B81634", # Red
    "#DF3B43", # Red orange
    "#E35C47", # Orange
    "#EB8749", # Light orange
    "#F6E254", # Yellow
    "#7B442D", # Brown
    "#000000", # Black
    "#1A6E7E", # Dark green - no longer dark green, now looks teal
    "#7CBF7B", # Green
    "#ADD2B8") # Light green
} else { # else (if black), remove black from colors
  tmp3 <- c(#"#EDEFEE", # Paper
    "#1A8BB3", # Teal - no longer teal, just bright blue
    "#0950AE", # Dark blue
    "#4DACE5", # Light blue
    "#126DDB", # Blue
    "#E48DC4", # Pink
    "#ABA9E8", # Light purple
    "#872791", # Purple
    "#6D1617", # Dark red
    "#B81634", # Red
    "#DF3B43", # Red orange
    "#E35C47", # Orange
    "#EB8749", # Light orange
    "#F6E254", # Yellow
    "#7B442D", # Brown
    #"#000000", # Black
    "#1A6E7E", # Dark green - no longer dark green, now looks teal
    "#7CBF7B", # Green
    "#ADD2B8") # Light green
}

# I think this is the orginial, plus a seed
func1 <- function(size, colors, seed=123){
  # set seed
  set.seed(seed)

  # Define an empty size x size matrix
  color_matrix <- matrix("", nrow = size, ncol = size)

  # For loop to go row by row through the empty matrix
  for (i in 1:nrow(color_matrix)){

    # For loop to go column by column
    for (j in 1:ncol(color_matrix)){

      # If you're in the first (top left) cell, assign a random color
      if (i == 1 && j == 1){

        color_matrix[i,j] <- sample(x = colors,
                                    size = 1,
                                    replace = TRUE)

        # If you're in any other cell than the top left
      } else {

        # Get the colors of the five surrounding cells, if they exist
        # (the condition in the ifelse checks to see if they exist)
        left_color1 <- ifelse((j-1) > 0, color_matrix[i, j-1], "")
        left_color2 <- ifelse((j-2) > 0, color_matrix[i, j-2], "")
        up_color1 <- ifelse((i-1) > 0, color_matrix[i-1, j], "")
        up_color2 <- ifelse((i-2) > 0, color_matrix[i-2, j], "")
        up_left_color <- ifelse((j-1) > 0 && (i-1) > 1, color_matrix[i-1, j-1], "")
        up_right_color <- ifelse((i-1) > 0 && (j+1) < (ncol(color_matrix)+1),
                                 color_matrix[i-1, j+1], "")

        # Check to see if any of the cell colors in cells that touch match, and
        # if they do, add the matching color to a vector of colors to be excluded
        # from the options when we assign a color to the cell we're on
        matching <- vector(mode = "character", length = 8)

        matching[1] <- ifelse(left_color1 == left_color2, left_color1, "")
        matching[2] <- ifelse(left_color1 == up_color1, left_color1, "")
        matching[3] <- ifelse(up_color1 == up_color2, up_color1, "")
        matching[4] <- ifelse(up_left_color == left_color1, up_left_color, "")
        matching[5] <- ifelse(up_left_color == up_color1, up_left_color, "")
        matching[6] <- ifelse(up_right_color == up_color1, up_right_color, "")
        matching[7] <- ifelse(up_left_color == left_color1, up_color1, "")
        matching[8] <- ifelse(up_left_color == up_color1, left_color1, "")

        # get just the unique colors in the list of matching colors,
        # ignoring any blank elements
        matching <- unique(matching[which(matching != "")])

        # If there were no matches
        if (length(matching) == 0){

          # Assign any random color
          color_matrix[i,j] <- sample(x = colors,
                                      size = 1,
                                      replace = TRUE)

        } else { # If there WERE matches

          # Assign any other color than those in the "matches" vector
          color_matrix[i,j] <- sample(x = colors[-which(colors %in% matching)],
                                      size = 1,
                                      replace = TRUE)
        }
      }
    }
  }
  # return the color matrix as a vector
  return(as.vector(color_matrix))
}

# I think this is just replacing ifelse and 1:nrow/ncol stuff
func2 <- function(size, colors, seed=123){
  # set seed
  set.seed(seed)

  # Define an empty size x size matrix
  color_matrix <- matrix("", nrow = size, ncol = size)

  # For loop to go row by row through the empty matrix
  for (i in seq_len(size)){

    # For loop to go column by column
    for (j in seq_len(size)){

      # If you're in the first (top left) cell, assign a random color
      if (i == 1 && j == 1){

        color_matrix[i,j] <- sample(x = colors,
                                    size = 1,
                                    replace = TRUE)

        # If you're in any other cell than the top left
      } else {

        # Get the colors of the five surrounding cells, if they exist
        # (the condition in the ifelse checks to see if they exist)
        left_color1 <- if((j-1) > 0) color_matrix[i, j-1] else ""
        left_color2 <- if((j-2) > 0) color_matrix[i, j-2] else ""
        up_color1 <- if((i-1) > 0) color_matrix[i-1, j] else ""
        up_color2 <- if((i-2) > 0) color_matrix[i-2, j] else ""
        up_left_color <- if((j-1) > 0 && (i-1) > 1) color_matrix[i-1, j-1] else ""
        up_right_color <- if((i-1) > 0 && (j+1) < (ncol(color_matrix)+1)) color_matrix[i-1, j+1] else ""

        # Check to see if any of the cell colors in cells that touch match, and
        # if they do, add the matching color to a vector of colors to be excluded
        # from the options when we assign a color to the cell we're on
        matching <- vector(mode = "character", length = 8)

        matching[1] <- if(left_color1 == left_color2) left_color1 else ""
        matching[2] <- if(left_color1 == up_color1) left_color1 else ""
        matching[3] <- if(up_color1 == up_color2) up_color1 else ""
        matching[4] <- if(up_left_color == left_color1) up_left_color else ""
        matching[5] <- if(up_left_color == up_color1) up_left_color else ""
        matching[6] <- if(up_right_color == up_color1) up_right_color else ""
        matching[7] <- if(up_left_color == left_color1) up_color1 else ""
        matching[8] <- if(up_left_color == up_color1) left_color1 else ""

        # get just the unique colors in the list of matching colors,
        # ignoring any blank elements
        matching <- unique(matching[which(matching != "")])

        # If there were no matches
        if (length(matching) == 0){

          # Assign any random color
          color_matrix[i,j] <- sample(x = colors,
                                      size = 1,
                                      replace = TRUE)

        } else { # If there WERE matches

          # Assign any other color than those in the "matches" vector
          color_matrix[i,j] <- sample(x = colors[-which(colors %in% matching)],
                                      size = 1,
                                      replace = TRUE)
        }
      }
    }
  }
  # return the color matrix as a vector
  return(as.vector(color_matrix))
}

# And then this one is the same as above, plus replacing the unique() call
func3 <- function(size, colors, seed=123){
  # set seed
  set.seed(seed)

  # Define an empty size x size matrix
  color_matrix <- matrix("", nrow = size, ncol = size)

  # For loop to go row by row through the empty matrix
  for (i in seq_len(size)){

    # For loop to go column by column
    for (j in seq_len(size)){

      # If you're in the first (top left) cell, assign a random color
      if (i == 1 && j == 1){

        color_matrix[i,j] <- sample(x = colors,
                                    size = 1,
                                    replace = FALSE)

        # If you're in any other cell than the top left
      } else {

        # Get the colors of the five surrounding cells, if they exist
        # (the condition in the ifelse checks to see if they exist)
        left_color1 <- if((j-1) > 0) color_matrix[i, j-1] else ""
        left_color2 <- if((j-2) > 0) color_matrix[i, j-2] else ""
        up_color1 <- if((i-1) > 0) color_matrix[i-1, j] else ""
        up_color2 <- if((i-2) > 0) color_matrix[i-2, j] else ""
        up_left_color <- if((j-1) > 0 & (i-1) > 1) color_matrix[i-1, j-1] else ""
        up_right_color <- if((i-1) > 0 & (j+1) < (ncol(color_matrix)+1)) color_matrix[i-1, j+1] else ""

        # Check to see if any of the cell colors in cells that touch match, and
        # if they do, add the matching color to a vector of colors to be excluded
        # from the options when we assign a color to the cell we're on
        matching <- vector(mode = "character", length = 8)

        matching[1] <- if(left_color1 == left_color2) left_color1 else ""
        matching[2] <- if(left_color1 == up_color1) left_color1 else ""
        matching[3] <- if(up_color1 == up_color2) up_color1 else ""
        matching[4] <- if(up_left_color == left_color1) up_left_color else ""
        matching[5] <- if(up_left_color == up_color1) up_left_color else ""
        matching[6] <- if(up_right_color == up_color1) up_right_color else ""
        matching[7] <- if(up_left_color == left_color1) up_color1 else ""
        matching[8] <- if(up_left_color == up_color1) left_color1 else ""

        # get just the unique colors in the list of matching colors,
        # ignoring any blank elements
        matching <- base::setdiff(matching, "")

        # If there were no matches
        color_matrix[i,j] <- sample(
          x = if (length(matching) == 0) colors else base::setdiff(colors, matching),
          size = 1,
          replace = FALSE)
        
      }
    }
  }
  # return the color matrix as a vector
  return(as.vector(color_matrix))
}



# speed test
mark(
  func1(tmp1, tmp3),
  func2(tmp1, tmp3),
  func3(tmp1, tmp3),
  iterations = 20,
  time_unit = "ns"
)
```

